---
title: "Project proposal"
author: "Red Team"
output: github_document
---

```{r load-packages, message = FALSE}
library(tidyverse)
library(broom)
```

## 1. Introduction

 Question: Which statistic in the NBA is most likely to correlate to MVP rank?
We will explore if there is a certain statistic in the NBA that has stronger relationship to rank than others. We will specifically look at points per game, team wins, and win shares. We predict that win shares will have the highest predictive power because it takes into consideration both individual player stats and how impact they are on their team. 

variables: rank, pts per game, team wins, win shares
cases: top 10 NBA players in MVP voting race from 1979-2018
data from: Github
https://github.com/dribbleanalytics/ml-mvp-predict/blob/master/2018-19-season/asb/final-csv-data/historical-mvps.csv


## 2. Data


2) (code in awards.Rmd)

388 rows, 27 columns

```{r data, eval=TRUE}
library(readr)
historical_mvps <- read_csv("proposal/historical-mvps.csv")
View(historical_mvps)

```




## 3. Data analysis plan

3)
Response: MVP voting rank
Explanatory: points per game, team wins, win shares

We want to make linear models for each explanatory variable and see which has a stronger relationship by looking at the r^2 value and the graph themselves. We will also combine variables together to see if that provides a stronger relationship. 


