---
title: "What stat is most useful in predicting MVP vote rankings?"
subtitle: "We will use data from 1979-2018 NBA seasons"
author: "Red Team <br> Naomi Korn, Stephen Chen, Zach Slavitt"
date: "`r Sys.Date()`"
output:
  xaringan::moon_reader:
    css: xaringan-themer.css
    lib_dir: libs
    nature:
      ratio: "16:9"
      highlightStyle: github
      highlightLines: true
      countIncrementalSlides: false
---

```{r load-packages, include = FALSE}
# Add any additional packages you need to this chunk

library(tidyverse)
library(tidymodels)
library(palmerpenguins)
library(knitr)
library(xaringanthemer)
library(showtext)
library(openintro)
```

```{r setup, include=FALSE}
# For better figure resolution
knitr::opts_chunk$set(fig.retina = 3, dpi = 300, fig.width = 6, fig.asp = 0.618, out.width = "80%")
```

```{r load-data, include=FALSE}
library(readr)
mvps <- read_csv("presentation_files/historical-mvps.csv")
view(mvps)

```

```{r, include=FALSE}
mvps <- mvps %>% 
  rename(
    team_wins = "Team Wins",
    pts_won = "Pts Won",
    overall_seed = "Overall Seed",
    fgpercent = "FG%"
  )
```

```{r, include=FALSE}
style_xaringan(
  title_slide_background_image = "img/red.jpg"
)
```


class: center, middle

## What stat has strongest relationship to MVP rank?

  
---

class: inverse, center, middle

# We will look at...
   ##1) Points per game
   ##2) Team wins
   ##3) Win Shares

---

# Rank versus Points per Game
.pull-left[

- **r^2 = 0.165** 

- slope = -0.247 

- weak correlation between increased rank and greater points per game

- Increased prevalence in 3-point shot could suggest that someone ranking 1 in 1979 might be scoring less points per game than someone ranking 10 in 2018. 
]
.pull-right[
```{r rankvspoints1, include=FALSE}
pts_fit <- lm(Rank ~ PTS, data = mvps)
tidy(pts_fit)
glance(pts_fit)
```
```{r rankvspointsgraph, warning=FALSE, out.width="80%", fig.width=4, echo=FALSE}
ggplot(data = mvps, aes(x = PTS, y = Rank)) +
  geom_jitter() +
   geom_smooth(method = "lm", se = FALSE, col="red")
```
]
---

class: inverse

# Rank versus Team Wins

.pull-left[
- **r^2 = 0.0983**

- Very weak correlation between rank and  team wins

- One explanation of this could be that players could be put in the top-10 just because they are on a good team, even though they have no chance of winning MVP
]
.pull-right[
```{r rankversusteam, include=FALSE}
rank_team_fit <- lm(Rank ~ team_wins, data = mvps)
tidy(rank_team_fit)
glance(rank_team_fit)
```

```{r teamwinsplot, warning=FALSE, out.width="90%", fig.width=4, echo=FALSE}
ggplot(data=mvps, mapping=aes(x=team_wins, y=Rank))+
geom_jitter()+
  geom_smooth(method="lm", se=FALSE)
```
]
---


# Rank versus Win Shares

.pull-left[
- "Win share: a measure that is assigned to players based on their offense, defense, and playing time. A win share is worth one-third of a team win. If a team wins 60 games, there are 180 'Win Shares' to distribute among the players."

- **r^2 = 0.3932**
- Stronger correlation between rank and win shares than the other statistics observed
]
.pull-right[

```{r rankvsWS, include=FALSE}
rank_ws_fit <- lm(Rank ~ WS, data = mvps)
summary(rank_ws_fit)
tidy(rank_ws_fit)
```

```{r wsplott, warning=FALSE, out.width="90%", fig.width=4, echo=FALSE}
ggplot(data=mvps, mapping=aes(x=WS, y=Rank))+
geom_jitter()+
  geom_smooth(method="lm", se=FALSE )
```
]


---

# Combining Variables

.pull-left[
```{r rankfit, include=FALSE}
rank_PTS_team_wins_fit <- lm(Rank ~ PTS + team_wins, data = mvps)
summary(rank_PTS_team_wins_fit)
```
```{r ptsandteamplot, warning=FALSE, out.width="90%", fig.width=4, echo=FALSE}
ggplot(data=mvps, mapping=aes(x= PTS + team_wins, y=Rank))+
geom_jitter()+
  geom_smooth(method="lm", se=FALSE )
```
```{r rankfit3stats, include=FALSE}
rank_PTS_team_WS_fit <- lm(Rank ~ PTS + team_wins + WS, data = mvps)
summary(rank_PTS_team_WS_fit)
glance(rank_PTS_team_WS_fit)
```
]
.pull-right[
```{r allvariableplot, warning=FALSE, out.width="90%", fig.width=4, echo=FALSE}
ggplot(data=mvps, mapping=aes(x=WS+team_wins+PTS, y=Rank))+
geom_jitter()+
  geom_smooth(method="lm", se=FALSE )
```
]


 ###The addition of WS into the prediction model strengthens the model. It increases the r^2 from 0.314 to 0.473.


---

# Conclusion

  ###WS is the best single predictor out of the 3 we examined
  ###The more variables, the better the model would be
  #Overall, no single stat determines who wins MVP, but rather an holistic view




