---
title: "Prediction of what stat is most useful in predicting MVP vote rankings"
subtitle: "Is there a one stat that is more predictive than others?"
author: "Red Team <br> Naomi Korn, Stephen Chen, Zach Slavitt"
date: "`r Sys.Date()`"
output:
  xaringan::moon_reader:
    css: xaringan-themer.css
    lib_dir: libs
    nature:
      ratio: "16:9"
      highlightStyle: github
      highlightLines: true
      countIncrementalSlides: false
---

```{r load-packages, include = FALSE}
# Add any additional packages you need to this chunk

library(tidyverse)
library(tidymodels)
library(palmerpenguins)
library(knitr)
library(xaringanthemer)
library(showtext)
library(openintro)
```

```{r setup, include=FALSE}
# For better figure resolution
knitr::opts_chunk$set(fig.retina = 3, dpi = 300, fig.width = 6, fig.asp = 0.618, out.width = "80%")
```

```{r load-data, include=FALSE}
library(readr)
mvps <- read_csv("presentation_files/historical-mvps.csv")
view(mvps)

```

```{r}
mvps <- mvps %>% 
  rename(
    team_wins = "Team Wins",
    pts_won = "Pts Won",
    overall_seed = "Overall Seed",
    fgpercent = "FG%"
  )
```
```{r}
onlymvps <- mvps %>%
  filter(Rank == 1)
view(onlymvps)
```
```{r}
nonmvps <- mvps %>%
  filter(Rank != 1)
view(nonmvps)
```
```{r}
style_xaringan(
  title_slide_background_image = "img/sports.jpg"
)
```


class: center, middle

## What is strongest predic
correlations
  points and MVP
  height and frequency
  age and frequency
  # of MVPs after winning 1 - what percentage of MVPs won one versus won multiple?
  age and MVP
  pick and MVP
  
---

class: inverse, center, middle

# Section title

---

# Hello World

- Click the `Knit` button to compile your presentation

- Make sure to commit and push all resulting files to your GitHub repo

---

class: inverse, middle, center

# Using xaringan

---

# xaringan

- The presentation is created using the `xaringan` package

- Use `---` to separate slides and `--` for incremental builds

--

- Like this

---

# Layouts

You can use plain text

- or bullet points

.pull-left[
or text in two columns $^*$
]
.pull-right[
- like
- this
]

.footnote[
[*] And add footnotes
]

---

# Code

```{r}
pts_fit <- linear_reg() %>%
  set_engine("lm") %>%
  fit(Rank ~ PTS, data = mvps)
```


```{r, eval=TRUE}
ggplot(mvps, aes(x = Rank)) +
  geom_density()
```
```{r point, eval=TRUE}
ggplot(mvps, aes(x = Rank, y = PTS)) +
  geom_jitter()
```
```{r profbeauty, eval=TRUE}
pts_fit <- lm(Rank ~ PTS, data = mvps)
tidy(pts_fit)
glance(pts_fit)
ggplot(data = mvps, aes(x = PTS, y = Rank)) +
  geom_jitter() +
   geom_smooth(method = "lm", se = FALSE, col="red")
```

```{r}
pts_aug <- augment(pts_fit$fit)
```

```{r}
names(pts_aug)
```
```{r residuals, eval=TRUE}
pts_aug %>%
ggplot(mapping = aes(x = .fitted, y = .resid)) +
  geom_jitter() +
 geom_hline(yintercept = 0)
```
```{r, eval=TRUE}
rank_team_fit <- lm(Rank ~ team_wins, data = mvps)
tidy(rank_team_fit)
glance(rank_team_fit)
```
```{r}
rank_ws_fit <- lm(Rank ~ WS, data = mvps)
summary(rank_ws_fit)
tidy(rank_ws_fit)
```

```{r genderfit, eval=TRUE}
rank_PTS_team_wins_fit <- lm(Rank ~ PTS + team_wins, data = mvps)
summary(rank_PTS_team_wins_fit)
```

```{r genderfit, eval=TRUE}
glance(rank_PTS_team_wins_fit)
r_value <- glance(rank_PTS_team_wins_fit)
select(r.squared)%>%
  pull()
r_value_percent <- r_value * 100
```
```{r}
rank_PTS_team_WS_fit <- lm(Rank ~ PTS + team_wins + WS, data = mvps)
summary(rank_PTS_team_WS_fit)
glance(rank_PTS_team_WS_fit)
```



```{r boring-regression3, eval=TRUE}
# a boring regression
ggplot(mvps, aes(x= PTS, y= team_wins))+
  geom_point()+
  geom_smooth(method = "lm")

```




```{r}
ggplot(mvps, aes(x=Rank, y=overall_seed)) +
  geom_jitter()
```

```{r}
ggplot(mvps, aes(x=overall_seed, fill= Rank)) +
  geom_bar()
```


---

# Plots


```{r positioncount, echo = FALSE}
ggplot(data=mvps, mapping=aes(x=team_wins, y=Rank))+
geom_jitter()+
  geom_smooth(method="lm")
```

```{r positioncount2, echo = FALSE}
ggplot(data=mvps, mapping=aes(x= PTS, y= Rank))+
geom_jitter()+
geom_smooth(method = "lm")
```


```{r positioncount5, echo = FALSE}
ggplot(data=mvps, mapping=aes(x= WS, y= Rank))+
geom_jitter()+
geom_smooth(method="lm")
```







```{r plot-iris, echo = FALSE}
# Code hidden with echo = FALSE
# Uses modified iris dataset from previous chunk
# Play around with height and width until you're happy with the look
ggplot(data = mvps, mapping = aes(x = team_wins, y = Age, color = Rank)) +
  geom_point() + 
  theme_minimal() # theme options: https://ggplot2.tidyverse.org/reference/ggtheme.html
```

---

## Plot and text

.pull-left[
- Some text
- goes here
]
.pull-right[
```{r warning=FALSE, out.width="100%", fig.width=4, echo=FALSE}
# see how I changed out.width and fig.width from defaults
# to make the figure bigger
ggplot(mvps, aes(x = team_wins, y = Age, color = Rank)) +
  geom_boxplot() +
  theme_minimal()
```
]

---

# Tables

If you want to generate a table, make sure it is in the HTML format (instead of Markdown or other formats), e.g.,

```{r iris-table, echo = FALSE}
kable(head(mvps), format = "html")
```

---

# Images

```{r castle, echo = FALSE, out.width = "60%", fig.align = "center", fig.cap = "Image credit: Photo by Jörg Angeli on Unsplash."}
include_graphics("https://images.unsplash.com/photo-1535448033526-c0e85c9e6968?ixlib=rb-1.2.1&ixid=eyJhcHBfaWQiOjEyMDd9&auto=format&fit=crop&w=1650&q=80")
```

Or you can also include a full page image. See next slide.

---

background-image: url(https://images.unsplash.com/photo-1535448033526-c0e85c9e6968?ixlib=rb-1.2.1&ixid=eyJhcHBfaWQiOjEyMDd9&auto=format&fit=crop&w=1650&q=80)

---

# Math Expressions

You can write LaTeX math expressions inside a pair of dollar signs, e.g. &#36;\alpha+\beta$ renders $\alpha+\beta$. You can use the display style with double dollar signs:

```
$$\bar{X}=\frac{1}{n}\sum_{i=1}^nX_i$$
```

$$\bar{X}=\frac{1}{n}\sum_{i=1}^nX_i$$

Limitations:

1. The source code of a LaTeX math expression must be in one line, unless it is inside a pair of double dollar signs, in which case the starting `$$` must appear in the very beginning of a line, followed immediately by a non-space character, and the ending `$$` must be at the end of a line, led by a non-space character;

1. There should not be spaces after the opening `$` or before the closing `$`.

1. Math does not work on the title slide (see [#61](https://github.com/yihui/xaringan/issues/61) for a workaround).

---

# Feeling adventurous?

- Want to find out more about `xaringan`? See https://slides.yihui.name/xaringan/#1.

- You are welcomed to use the default styling of the slides. In fact, that's what I expect majority of you will do. You will differentiate yourself with the content of your presentation.

- But some of you might want to play around with slide styling. The 
`xaringanthemer` provides some solutions for this that: https://pkg.garrickadenbuie.com/xaringanthemer.

- And if you want more bells and whistles, there is also `xaringanExtra`: https://pkg.garrickadenbuie.com/xaringanExtra.
